# Changing rasta size of climate data for iLand
Goal is to change the current raster size from 14x14km down to 1x1km and link the specific climates to this. For this rasta data is used and linked to climate data from the dwd. Afterwards a cluster analysis is done to identify clusters and combine the climate to the center of each clusters.
iLand needs the following climate columns:
- point_id
- year
- month
- day
- min_temp
- max_temp
- prec
- rad
- vpd

Workflow should be in the following steps:
- gather relevant data
- cluster analysis
- z-skaling of climate data
- Get data for the center of the clusters
```{r}
if (.Platform$OS.type == "windows") {
  cPath <- "F:/data/climatedata/germany_1km_res/2_work"
  cRawPath <- "F:/data/climatedata/germany_1km_res/1_dataRaw"
} else {
  cPath <- "/data/public/data/climatedata/germany_1km_res/2_work"
  cRawPath <- "/data/public/data/climatedata/germany_1km_res/1_dataRaw"
}
cPath <- normalizePath(cPath, winslash = "/")
cRawPath <- normalizePath(cRawPath, winslash = "/")

gauss_kruger_crs <- "EPSG:31467"
```
Loading Libraries that are used throughout this script.
```{r}
library(terra)
library(raster)
library(tidyverse)
library(weathermetrics)
library(factoextra)
library(cluster)
library(ggpubr)
library(sf)
#library(compiler)
#library(fastkmedoids)
```
## Gather climate data
Load raster data.
```{r}
setwd(cPath)
ann_max_tmp <- terra::rast(file.path(cRawPath,"raster","TADXMM_17_1981_30.asc"))
ann_min_tmp <- terra::rast(file.path(cRawPath,"raster","TADNMM_17_1981_30.asc"))
ann_mean_prec <- terra::rast(file.path(cRawPath,"raster","RSMS_17_1981_30.asc"))
```
Next is extracting the information stored in the raster data and store it in a table. First a function due to repeated process.
```{r}
get_raster_data <- function(rRaster) {
  coords <- terra::xyFromCell(rRaster, 1:ncell(rRaster))
  vals <- terra::extract(rRaster, coords)
  r_df <- data.frame(
    lon = coords[,1],
    lat = coords[,2],
    val = vals[,1]
  )
  return(r_df)
}
```
Extracting data from raster into table, for precipitation, min & max temperature
```{r}
df_prec <- get_raster_data(ann_mean_prec) %>% rename(prec = val)
df_max_tmp <- get_raster_data(ann_max_tmp) %>% rename(max_tmp = val)
df_min_tmp <- get_raster_data(ann_min_tmp) %>% rename(min_tmp = val)
df_clim <- inner_join(
  x = df_max_tmp,
  y = df_min_tmp,
  by = join_by(lat, lon)
) %>% 
  inner_join(
    x = .,
    y = df_prec,
    by = join_by(lat,lon)
  )

rm(df_prec, df_max_tmp, df_min_tmp)
```
### More complex .asc files
Some of the DWD files contain a custom header. The terra package cannot handle this, so a custom approach is needed. Here is a function to get the rasta data from a .asc file with a custom header, it will skip the header and extract only the "raster" part of the file.
```{r}
get_raster <- function(strFilePath) {
  # Read the full file as text
  # lines <- readLines(strFilePath, encoding = "UTF-8")
  lines <- readLines(strFilePath, encoding = "latin1")
  
  # Find the line where raster data begins 
  # (the "[ASCII-Raster-Format]" block starts here)
  start_idx <- grep("^NCOLS", lines)  # or grep("^[A-Z]+", lines) to find NCOLS
  
  # Keep only the raster part
  ascii_raster_lines <- lines[start_idx:length(lines)]

  # Write it to a unique temporary file
  # writeLines(ascii_raster_lines, file.path(cPath,"raster_cleared.asc"))
  tmp_file <- tempfile(fileext = ".asc")
  writeLines(ascii_raster_lines, tmp_file)
  
  # # Now load it with terra
  # r <- terra::rast(file.path(cPath,"raster_cleared.asc"))
  r <- terra::rast(tmp_file)
  
  # Try without deleting manually
  # # Delete temporary file
  # unlink(tmp_file)
  cat("Raster: ", summary(r, size = Inf), "\n"); flush.console()

  return(r)
}
```
Extract radiation data from raster.
```{r}
# Create raster from multi monthly radiation global for all 12 months
rad_months <- list.files(
  file.path(cRawPath, "raster", "20250714"),
  full.names = T,
  pattern = "monthly_rad"
)
rad_months_rast <- rast(lapply(X = rad_months, FUN = get_raster))
ann_rad <- app(rad_months_rast, "sum")
```
### Calculating vapour pressure deficite (VPD)
```{r}
vpd_calc <- function(t, td, c1 = 0.611, c2 = 17.67, c3 = 243.5) {
  c1 * exp((c2 * (t)) / (t + c3)) - c1 * exp((c2 * (td)) / (td + c3))
}
```
Loading hurs (humidity) and tas (mean temperature) rasters form DWD into variables and calculating VPD from these variables.
```{r}
# Loading hurs & tas raster
eval_hurs <- rast(
  file.path(cRawPath,"raster","hurs_hyras_1_1981_2010_v6-0_de_mean.nc")
)
plot(eval_hurs)
eval_tas <- rast(
  file.path(cRawPath,"raster","tas_hyras_1_1981_2010_v6-0_de_mean.nc")
)
plot(eval_tas)

# Calculating VPD
td <- weathermetrics::humidity.to.dewpoint(
  rh = eval_hurs,
  t = eval_tas,
  temperature.metric = "celsius"
)

eval_vpd <- vpd_calc(eval_tas, td)
plot(eval_vpd)

# Extracting the information form the raster and add it to the climate frame
df_vpd <- get_raster_data(eval_vpd)

```
### Combining climate data
Combining all climate dataframes into one. Here the different extents will need to be harmonised.
```{r}
# min, max, prec, rad are in gauss-krueger
gauss_kruger_crs <- "EPSG:31467"
crs(ann_max_tmp) <- gauss_kruger_crs
crs(ann_min_tmp) <- gauss_kruger_crs
crs(ann_mean_prec) <- gauss_kruger_crs
crs(ann_rad) <- gauss_kruger_crs

# vpd is in ETRS89 / LAEA Europe
crs(eval_vpd)

# converting etr back to gk
vpd_pr <- project(eval_vpd, crs(ann_max_tmp))

# resampling all raster to the same extend
vpd_res <- resample(vpd_pr, ann_max_tmp, method = "bilinear")
rad_res <- terra::resample(ann_rad, ann_max_tmp, method = "bilinear")
plot(ann_max_tmp)
plot(ann_mean_prec)
plot(vpd_res)
plot(rad_res)

# Get data from raster
df_rad <- get_raster_data(rad_res)
df_vpd <- get_raster_data(vpd_res)

# Add rad and vpd to df_clim
df_clim <- inner_join(
  x = df_clim,
  y = df_rad,
  by = join_by(lat, lon)
) %>% inner_join(
  x = .,
  y = df_vpd,
  by = join_by(lat, lon)
)
df_clim <- df_clim %>% rename(rad = val.x, vpd = val.y)
df_clim$max_tmp <- df_clim$max_tmp /10 # Scaling to °C
df_clim$min_tmp <- df_clim$min_tmp /10 # Scaling to °C
df_clim$rad <- df_clim$rad * 3.6 # Scaling from kWh/m² to MJ/m²/day
```
Cleanup before cluster analysis:
```{r}
rm(ann_max_tmp, ann_mean_prec, ann_min_tmp, ann_rad)
rm(eval_hurs, eval_tas, eval_vpd, df_rad, df_vpd)
rm(td, vpd_pr, vpd_res, rad_res)

# write_csv(df_clim %>% na.omit(), file = "clim_dat.csv")
```

#### f_create_grid
Create function to convert a data frame to a geographical object.

```{r}
f_create_grid <- function(df, refsystem, cell_size = 1000) {
  # Create 1km buffer squares (0.5km radius in each direction)
  l_sf <- st_as_sf(
    x = df,
    coords = c("lon", "lat"),
    crs = refsystem
  )

  # Reproject to meters (for buffering), then buffer, then back to GK
  l_grid <- l_sf %>%
    st_buffer(
      dist = cell_size / 2,
      endCapStyle = "SQUARE"
    ) # make square polygons

  # Return grid
  return(l_grid)
}
```
