# Post clustering processing
Here the clustered climate data is processed, meaning the deviation from each point towards the centoid is calculated and a representative datapoint for each cluster is chosen. In the end a .csv file is created containing the information about the cluster center and the representative plot.
```{r}
# Libraries
library(dplyr)
library(tidyr)
library(ggpubr)
library(readr)
```
## 1.  Set path to working environment independent of operating system
```{r}
if (.Platform$OS.type == "windows") {
  cPath <- "F:/Data/climatedata/germany_1km_res/2_work"
} else {
  cPath <- "/data/public/Data/climatedata/germany_1km_res/2_work"
}
cPath <- normalizePath(cPath, winslash = "/")
```
## 2.  Load csv files containing information about the clusters and the climate data
```{r}
setwd(cPath)
# Clustered climate
dfClimDat <- read.csv(file = file.path("clustered","k10000_clim_dat.csv"))
# Cluster center
dfCentoid <- read.csv(file = file.path("clustered", "k10000_centoids.csv"))
```
## 3.  Search closest
```{r}
r_df <- data.frame(
  cluster = character(length(unique(dfClimDat$cluster))),
  closestP = character(length(unique(dfClimDat$cluster)))
)
i <- 1
for (icluster in unique(dfClimDat$cluster)) {
  df1 <- dfCentoid[dfCentoid$cluster_id == icluster,] %>% dplyr::select(-rep_index)
  df2 <- dfClimDat[dfClimDat$cluster == icluster,] %>% dplyr::select(-lat, -lon)
  
  df1 <- df1 %>% dplyr::select(-cluster_id)
  df2 <- df2 %>% dplyr::select(-cluster)
  
  idist <- which.min(sqrt(rowSums(df2 - df1[rep(1, nrow(df2)),])^2))
  
  r_df[i, 1] <- icluster
  r_df[i, 2] <- names(idist)
  i <- i + 1
}
rm(df1, df2, i, icluster, idist)
```
## 4.  Match cluster center point with latitude and longitude
```{r}
dfClimDat$id <- rownames(dfClimDat)
dfClimCent <- dfClimDat[dfClimDat$id %in% r_df$closestP,]
rm(r_df)
```
## 5.  Calculate deviation from cluster centoid
```{r}
dfCombi <- left_join(
  x = dfClimDat,
  y = dfClimCent %>% dplyr::select(-c(lat, lon, id)),
  by = join_by(cluster),
  suffix = c("","_cent")
) %>% mutate(
  max_tmp_dif = max_tmp - max_tmp_cent,
  min_tmp_dif = min_tmp - min_tmp_cent,
  prec_dif = prec - prec_cent,
  rad_dif = rad - rad_cent,
  vpd_dif = vpd - vpd_cent
)
dfClDif <- left_join(
  x = dfClimDat,
  y = dfCentoid %>% dplyr::select(-rep_index),
  by = join_by(cluster == cluster_id),
  suffix = c("","_cent")
) %>% mutate(
  max_tmp_dif = max_tmp - max_tmp_cent,
  min_tmp_dif = min_tmp - min_tmp_cent,
  prec_dif = prec - prec_cent,
  rad_dif = rad - rad_cent,
  vpd_dif = vpd - vpd_cent
)
```
Some cool plots to show the differences between centoid and each row.
```{r}
ggplot(data = dfCombi %>% pivot_longer(
  cols = c(ends_with("_dif")),
  names_to = "difs",
  values_to = "dif_val"
)) +
  geom_density(
    aes(x = dif_val),
    alpha = .5
  ) +
  facet_wrap(~difs, scales = "free") +
  labs(title = "Difference to centoid") +
  theme_pubr()

ggplot(data = dfClDif %>% pivot_longer(
  cols = c(ends_with("_dif")),
  names_to = "difs",
  values_to = "dif_val"
)) +
  geom_density(
    aes(x = dif_val),
    alpha = .5
  ) +
  facet_wrap(~difs, scales = "free") +
  labs(title = "Difference to cluster mean") +
  theme_pubr()
```
## 6.  Safe .csv file with centoid latitude and longitude
```{r}
write_csv(
  #dfCombi %>% select(lat, lon, cluster),
  dfClimCent %>% dplyr::select(lon, lat, cluster),
  file = "cluster_coords.csv"
)
```

Cleanup:
```{r}
rm(cPath)
rm(dfCentoid, dfClimCent)
rm(dfCombi, dfClimDat, dfClDif)
```
