# Work with pyton .csv (post-cluster process)
BEFORE CONTINUING, MAKE SURE THE "02_cluster5-10k.py" WAS RUN SUCCESSFULLY!
Clustering was done via a python script, so the clusters for 5-10k are loaded from there.
```{r}
if (.Platform$OS.type == "windows") {
  cPath <- "F:/data/climatedata/germany_1km_res/2_work"
  # cRawPath <- "F:/data/climatedata/germany_1km_res/1_dataRaw"
} else {
  cPath <- "/data/public/data/climatedata/germany_1km_res/2_work"
  # cRawPath <- "/data/public/data/climatedata/germany_1km_res/1_dataRaw"
}
cPath <- normalizePath(cPath, winslash = "/")
# cRawPath <- normalizePath(cRawPath, winslash = "/")

gauss_kruger_crs <- "EPSG:31467"
```
Loading libraries
```{r}
library(terra)
library(tidyverse)
library(ggpubr)
library(patchwork)
library(flextable)
library(sf)
```
Load clustered climate data from .csv files.
```{r}
setwd(cPath)
df_5k <- read.csv(file = file.path("clustered","k5000_clim_dat.csv"))
df_6k <- read.csv(file = file.path("clustered","k6000_clim_dat.csv"))
df_7k <- read.csv(file = file.path("clustered","k7000_clim_dat.csv"))
df_8k <- read.csv(file = file.path("clustered","k8000_clim_dat.csv"))
df_9k <- read.csv(file = file.path("clustered","k9000_clim_dat.csv"))
df_10k <- read.csv(file = file.path("clustered","k10000_clim_dat.csv"))
df_15k <- read.csv(file = file.path("clustered","k15000_clim_dat.csv"))
```
#### Calculate Anomalies
##### Anomalies from cluster mean
```{r}
f_calcAnom <- function(dfclim) {
  # Calculate cluster means for each variable
  cluster_means <- dfclim %>%
    group_by(cluster) %>%
    summarise(
      across(
        c(max_tmp, min_tmp, prec, rad, vpd),
        mean,
        .names = "mean_{.col}"
      )
    )
  
  # Join the means back to the original data
  climate_anomalies <- dfclim %>%
    left_join(cluster_means, by = "cluster") %>%
    mutate(
      max_temp_anomaly = max_tmp - mean_max_tmp,
      min_temp_anomaly = min_tmp - mean_min_tmp,
      prec_anomaly = prec - mean_prec,
      rad_anomaly = rad - mean_rad,
      vpd_anomaly = vpd - mean_vpd
    )
  return(climate_anomalies)
}

f_create_grid <- function(df, refsystem, cell_size = 1000) {
  # Create 1km buffer squares (0.5km radius in each direction)
  l_sf <- st_as_sf(
    x = df,
    coords = c("lon", "lat"),
    crs = refsystem
  )

  # Reproject to meters (for buffering), then buffer, then back to GK
  l_grid <- l_sf %>%
    st_buffer(
      dist = cell_size / 2,
      endCapStyle = "SQUARE"
    ) # make square polygons

  # Return grid
  return(l_grid)
}
```
##### Anomalies from cluster medoid
Load centoids of clusters
```{r}
# .csv with centoids are saved next to the clustered climate data
df15k_cent <- read.csv(file = file.path("clustered", "k15000_centoids.csv"))
df15k_cent <- df15k_cent %>% rename(cluster = cluster_id)

df5k_cent <- read.csv(file = file.path("clustered", "k5000_centoids.csv"))
df6k_cent <- read.csv(file = file.path("clustered", "k6000_centoids.csv"))
df7k_cent <- read.csv(file = file.path("clustered", "k7000_centoids.csv"))
df8k_cent <- read.csv(file = file.path("clustered", "k8000_centoids.csv"))
df9k_cent <- read.csv(file = file.path("clustered", "k9000_centoids.csv"))
df10k_cent <- read.csv(file = file.path("clustered", "k10000_centoids.csv"))

df5k_cent <- df5k_cent %>% rename(cluster = cluster_id)
df6k_cent <- df6k_cent %>% rename(cluster = cluster_id)
df7k_cent <- df7k_cent %>% rename(cluster = cluster_id)
df8k_cent <- df8k_cent %>% rename(cluster = cluster_id)
df9k_cent <- df9k_cent %>% rename(cluster = cluster_id)
df10k_cent <- df10k_cent %>% rename(cluster = cluster_id)
```

Function to claculate the anomalies
```{r}
f_calcCentAnom <- function(df, df_centoids) {
  climate_anomalies <- left_join(
    x = df,
    y = df_centoids,
    by = "cluster",
    suffix = c("", "_cent")
  ) %>% 
    mutate(
      max_temp_anomaly = max_tmp - max_tmp_cent,
      min_temp_anomaly = min_tmp - min_tmp_cent,
      prec_anomaly = prec - prec_cent,
      rad_anomaly = rad - rad_cent,
      vpd_anomaly = vpd - vpd_cent
    )
  return(climate_anomalies)
}
```

#### cont. Anomalie calculation
Identify closest point to the centoid for each cluster
```{r}
df_5k <- f_calcCentAnom(df_5k, df5k_cent)
df_6k <- f_calcCentAnom(df_6k, df6k_cent)
df_7k <- f_calcCentAnom(df_7k, df7k_cent)
df_8k <- f_calcCentAnom(df_8k, df8k_cent)
df_9k <- f_calcCentAnom(df_9k, df9k_cent)
df_10k <- f_calcCentAnom(df_10k, df10k_cent)
df_15k <- f_calcCentAnom(df_15k, df15k_cent)
```
## Anomaly histograms
### Density - max_temp
```{r}
ggplot() +
  theme_pubr() +
  geom_density(
    data = df_5k,
    aes(x = max_temp_anomaly, col = "5k"),
    alpha = .5
  ) +
  geom_density(
    data = df_6k,
    aes(x = max_temp_anomaly, col = "6k"),
    alpha = .5
  ) +
  geom_density(
    data = df_7k,
    aes(x = max_temp_anomaly, col = "7k"),
    alpha = .5
  ) +
  geom_density(
    data = df_8k,
    aes(x = max_temp_anomaly, col = "8k"),
    alpha = .5
  ) +
  geom_density(
    data = df_9k,
    aes(x = max_temp_anomaly, col = "9k"),
    alpha = .5
  ) +
  geom_density(
    data = df_10k,
    aes(x = max_temp_anomaly, col = "10k"),
    alpha = .5
  ) +
  geom_density(
    data = df_15k,
    aes(x = max_temp_anomaly, col = "15k")
  ) +
  #xlim(-2.5,2.5) +
  scale_color_manual(
    name="Cluster sizes",
    breaks=c("5k", "6k", "7k", "8k", "9k", "10k", "15k"),
    values=c("5k" = "#bfa19a", "6k" = "#62a2b3",
             "7k" = "#7a5048", "8k" = "#452e2a",
             "9k" = "#cd6548", "10k" = "#2d9eb6",
             "15k" = "#0f0a0a")
  )
```
### Density - min_temp
```{r}
ggplot() +
  theme_pubr() +
  geom_density(
    data = df_5k,
    aes(x = min_temp_anomaly, col = "5k"),
    alpha = .5
  ) +
  geom_density(
    data = df_6k,
    aes(x = min_temp_anomaly, col = "6k"),
    alpha = .5
  ) +
  geom_density(
    data = df_7k,
    aes(x = min_temp_anomaly, col = "7k"),
    alpha = .5
  ) +
  geom_density(
    data = df_8k,
    aes(x = min_temp_anomaly, col = "8k"),
    alpha = .5
  ) +
  geom_density(
    data = df_9k,
    aes(x = min_temp_anomaly, col = "9k"),
    alpha = .5
  ) +
  geom_density(
    data = df_10k,
    aes(x = min_temp_anomaly, col = "10k"),
    alpha = .5
  ) +
  geom_density(
    data = df_15k,
    aes(x = min_temp_anomaly, col = "15k")
  ) +
  #xlim(-2.5,2.5) +
  scale_color_manual(
    name="Cluster sizes",
    breaks=c("5k", "6k", "7k", "8k", "9k", "10k", "15k"),
    values=c("5k" = "blue", "6k" = "yellow3",
             "7k" = "orange", "8k" = "grey",
             "9k" = "red2", "10k" = "black", "15k" = "green")
  )
```
### Density - prec
```{r}
ggplot() +
  theme_pubr() +
  geom_density(
    data = df_5k,
    aes(x = prec_anomaly, col = "5k"),
    alpha = .5
  ) +
  geom_density(
    data = df_6k,
    aes(x = prec_anomaly, col = "6k"),
    alpha = .5
  ) +
  geom_density(
    data = df_7k,
    aes(x = prec_anomaly, col = "7k"),
    alpha = .5
  ) +
  geom_density(
    data = df_8k,
    aes(x = prec_anomaly, col = "8k"),
    alpha = .5
  ) +
  geom_density(
    data = df_9k,
    aes(x = prec_anomaly, col = "9k"),
    alpha = .5
  ) +
  geom_density(
    data = df_10k,
    aes(x = prec_anomaly, col = "10k"),
    alpha = .5
  ) +
  geom_density(
    data = df_15k,
    aes(x = prec_anomaly, col = "15k"),
    alpha = .5
  ) +
  #xlim(-100,100) +
  scale_color_manual(
    name="Cluster sizes",
    breaks=c("5k", "6k", "7k", "8k", "9k", "10k", "15k"),
    values=c("5k" = "blue", "6k" = "yellow3",
             "7k" = "orange", "8k" = "grey",
             "9k" = "red2", "10k" = "black",
             "15k" = "green")
  )
```
### Density - rad
```{r}
ggplot() +
  theme_pubr() +
  geom_density(
    data = df_5k,
    aes(x = rad_anomaly, col = "5k"),
    alpha = .5
  ) +
  geom_density(
    data = df_6k,
    aes(x = rad_anomaly, col = "6k"),
    alpha = .5
  ) +
  geom_density(
    data = df_7k,
    aes(x = rad_anomaly, col = "7k"),
    alpha = .5
  ) +
  geom_density(
    data = df_8k,
    aes(x = rad_anomaly, col = "8k"),
    alpha = .5
  ) +
  geom_density(
    data = df_9k,
    aes(x = rad_anomaly, col = "9k"),
    alpha = .5
  ) +
  geom_density(
    data = df_10k,
    aes(x = rad_anomaly, col = "10k"),
    alpha = .5
  ) +
  geom_density(
    data = df_15k,
    aes(x = rad_anomaly, col = "15k"),
    alpha = .5
  ) +
  #xlim(-1.5,1.5) +
  scale_color_manual(
    name="Cluster sizes",
    breaks=c("5k", "6k", "7k", "8k", "9k", "10k", "15k"),
    values=c("5k" = "blue", "6k" = "yellow3",
             "7k" = "orange", "8k" = "grey",
             "9k" = "red2", "10k" = "black",
             "15k" = "green")
  )
```

### Density - vpd
```{r}
ggplot() +
  theme_pubr() +
  geom_density(
    data = df_5k,
    aes(x = vpd_anomaly, col = "5k"),
    alpha = .5
  ) +
  geom_density(
    data = df_6k,
    aes(x = vpd_anomaly, col = "6k"),
    alpha = .5
  ) +
  geom_density(
    data = df_7k,
    aes(x = vpd_anomaly, col = "7k"),
    alpha = .5
  ) +
  geom_density(
    data = df_8k,
    aes(x = vpd_anomaly, col = "8k"),
    alpha = .5
  ) +
  geom_density(
    data = df_9k,
    aes(x = vpd_anomaly, col = "9k"),
    alpha = .5
  ) +
  geom_density(
    data = df_10k,
    aes(x = vpd_anomaly, col = "10k"),
    alpha = .5
  ) +
  geom_density(
    data = df_15k,
    aes(x = vpd_anomaly, col = "15k"),
    alpha = .5
  ) +
  #xlim(-0.02,0.02) +
  scale_color_manual(
    name="Cluster sizes",
    breaks=c("5k", "6k", "7k", "8k", "9k", "10k", "15k"),
    values=c("5k" = "blue", "6k" = "yellow3",
             "7k" = "orange", "8k" = "grey",
             "9k" = "red2", "10k" = "black",
             "15k" = "green")
  )
```
### Metrics for Anomalies (table 2)
Calculate area mean and sd
```{r}
f_process_area <- function(df, crs_refsystem) {
  # df %>%
  #   f_create_grid(refsystem = crs_refsystem) %>%
  #   group_by(cluster) %>%
  #   summarise(
  #     uni = st_union(geometry),
  #     area_km2 = round(as.numeric(st_area(uni)) / 1e6, 3),
  #     .groups = "drop"
  #   ) %>%
  #   mutate(
  #     area_km2_mean = mean(area_km2),
  #     area_km2_sd   = sd(area_km2)
  #   )
  
  
  cluster_areas <- df %>%
    f_create_grid(refsystem = crs_refsystem) %>%
    select(cluster, geometry) %>% 
    group_by(cluster) %>%
    summarise(
      geom = st_combine(geometry),
      area_km2 = as.numeric(st_area(geom)) / 1e6,
      .groups = "drop"
    ) %>% 
    summarise(
      area_km2_mean = mean(area_km2),
      area_km2_sd   = sd(area_km2)
    )
  return(cluster_areas)

}
# df_area <- df_list %>%
#   map(f_process_area, crs_refsystem = crs_refsystem) %>%
#   list_rbind(names_to = "source_id")

library(furrr)

plan(multicore)  # or multicore on Linux/macOS

df_area <- df_list %>%
  future_map(f_process_area, crs_refsystem = crs_refsystem) %>%
  list_rbind(names_to = "source_id")

df_area <- df_area %>% mutate(
  cluster = str_remove(source_id, "cluster")
) %>% 
  select(
    -c(source_id, geom)
  )

```

Calculate climatic means and sd
```{r}
# ---- STEP 1: Put dataframes into a list ----
df_list <- list(
  cluster5k = df_5k,
  cluster6k = df_6k,
  cluster7k = df_7k,
  cluster8k = df_8k,
  cluster9k = df_9k,
  cluster10k = df_10k,
  cluster15k = df_15k
)

# ---- STEP 2: Specify parameter columns ----
parameter_cols <- c("max_temp_anomaly", "min_temp_anomaly", "prec_anomaly", "rad_anomaly", "vpd_anomaly")

# ---- STEP 3: Function to compute mean ± sd for one dataframe ----
f_summarize_mean_sd <- function(df, cols) {
  r_df <- df %>%
    summarise(
      across(
        all_of(cols),
        ~ sprintf("%.3f ± %.3f", mean(.x, na.rm = TRUE), sd(.x, na.rm = TRUE))
      )
    )
  return(r_df)
}

# ---- STEP 4: Apply summarization to all dataframes ----
summary_table <- bind_rows(
  map(df_list, f_summarize_mean_sd, cols = parameter_cols),
  .id = "cluster"
) %>% mutate(
  cluster = str_remove(cluster, "cluster"),
  storage = c("95.5", "115", "134", "153", "172", "191", "287")
) %>% rename(
  Cluster = cluster,
  `Residual multi-annual maximum temperature [°C]` = max_temp_anomaly,
  `Residual multi-annual minimum temperature [°C]` = min_temp_anomaly,
  `Residual multi-annual precipitation [mm]` = prec_anomaly,
  `Residual multi-annual global radiation [MJm-2day-1]` = rad_anomaly,
  `Residual multi-annual vapour pressure deficit [kPa]` = vpd_anomaly,
  `Storage requirement [GB]` = storage
)

summary_table <- left_join(
  x = summary_table,
  y = df_area %>% 
    mutate(
      `Mean cluster area [km²]` = sprintf("%.3f ± %.3f", area_km2_mean, area_km2_sd)
    ) %>% select(-c(area_km2_mean, area_km2_sd)),
  by = join_by(Cluster == cluster)
)


ft <- flextable(summary_table) %>%
  #merge_v(j = "ecosystem") %>%
  theme_booktabs() %>%
  color(i = seq(2, nrow(summary_table), 2), color = "black", part = "body") %>%
  bg(i = seq(2, nrow(summary_table), 2), bg = "#F9F9F9", part = "body") %>%
  autofit() %>%
  align(align = "center", part = "all") %>%
  bold(part = "header")
ft
ft <- fit_to_width(ft, max_width = 6.5)
flextable::save_as_docx(ft, path = file.path("output", "table2.docx"))
```

## Plot Map
```{r}
crs_refsystem <- 31467

df5k_sf <- f_create_grid(
  df = df_5k,
  refsystem = crs_refsystem
)
df10k_sf <- f_create_grid(
  df = df_10k,
  refsystem = crs_refsystem
)
df15k_sf <- f_create_grid(
  df = df_15k,
  refsystem = crs_refsystem
)
```

```{r}
ggplot(data = df10k_sf) +
  geom_sf(aes(color = max_temp_anomaly), size = 1) +
  scale_color_gradient2(
    low = "blue",
    mid = "white",
    high = "red",
    midpoint = 0,
    limit = c(-2,2)
  ) +
  theme_minimal() +
  labs(
    title = "Max. Temperature Anomaly from Cluster Mean -10k",
    color = "Anomaly (°C)"
  )

ggplot(data = df10k_sf) +
  geom_sf(aes(color = prec_anomaly), size = 1) +
  scale_color_gradient2(
    low = "blue",
    mid = "white",
    high = "red",
    midpoint = 0,
    limit = c(-2,2)
  ) +
  theme_minimal() +
  labs(
    title = "Precipitation anomaly from cluster mean -10k",
    color = "Anomaly (mm)"
  )

ggplot(data = df10k_sf) +
  geom_sf(aes(color = prec), size = 1) +
  theme_minimal()
```
### Create tiff with clusters
```{r}
# Create SpatRaster
rast_15k <- terra::rast(
  df_15k, # %>% select(lon, lat, cluster),
  type = "xyz",
  crs = gauss_kruger_crs
)
# Plot to check
plot(rast_15k[["cluster"]])

rast_5k <- terra::rast(
  df_5k,
  type = "xyz",
  crs = gauss_kruger_crs
)
plot(rast_5k[["cluster"]])

rast_10k <- terra::rast(
  df_10k,
  type = "xyz",
  crs = gauss_kruger_crs
)
plot(rast_10k[["cluster"]])
```

```{r}
# write .tiff file
# terra::writeRaster(
#   x = rast_15k,
#   filename = file.path("tif","15k_clim_cl.tif"),
#   overwrite = TRUE
# )
# terra::writeRaster(
#   x = rast_5k,
#   filename = file.path("tif","5k_clim_cl.tif"),
#   overwrite = TRUE
# )
# terra::writeRaster(
#   x = rast_10k,
#   filename = file.path("tif","10k_clim_cl.tif"),
#   overwrite = TRUE
# )
```

For getting climate data on the 1km² scale, the 10k clusters will be used.

# Creating reference raster
Creating a raster that contains the coordinates and the associate cluster of every km² in Germany.
```{r}
df10k_clust <- read_csv(file = "clustered/k10000_clim_dat.csv")
#ref_rast <- f_create_grid(df = df10k_clust, refsystem = crs_refsystem)

ref_rast <- terra::rast(
  df10k_clust,
  type = "xyz",
  crs = gauss_kruger_crs
)
plot(ref_rast[["prec"]])
plot(ref_rast[["cluster"]])

# terra::writeRaster(
#   x = ref_rast,
#   filename = "tif/reference_raster.tif"
# )
```

# Paper - Figures
First create a dataframe for all cluster scenarios.
```{r}
df_5k$nr_cl <- as.factor(5000)
df_6k$nr_cl <- as.factor(6000)
df_7k$nr_cl <- as.factor(7000)
df_8k$nr_cl <- as.factor(8000)
df_9k$nr_cl <- as.factor(9000)
df_10k$nr_cl <- as.factor(10000)
df_15k$nr_cl <- as.factor(15000)

df_clScen <- bind_rows(
  df_5k,
  df_6k,
  df_7k,
  df_8k,
  df_9k,
  df_10k,
  df_15k
)

ldf_clScen <- df_clScen %>% pivot_longer(
  cols = ends_with("_anomaly"),
  names_to = "anomaly_param",
  values_to = "anomaly_val"
)

cfacet_names <- c(
  "max_temp_anomaly" = "max. temperature",
  "min_temp_anomaly" = "min. temperature",
  "prec_anomaly" = "precipitation",
  "rad_anomaly" = "rad",
  "vpd_anomaly" = "vpd"
)
```
## Fig. 2. Cluster maps and densitys plots
Plot cluster scenarios with density plot for anomalies.
```{r}
p_dens <- ggplot(data = ldf_clScen) +
  geom_vline(linetype = 4, xintercept = 0) +
  geom_freqpoly(
    aes(
      x = anomaly_val,
      colour = nr_cl
    ),
    bins = 100,
    linewidth = 0.8,
    alpha = 0.8
  ) +
  facet_wrap(
    ~ anomaly_param,
    ncol = 1,
    nrow = 5,
    scales = "free",
    labeller = labeller(anomaly_param = cfacet_names)
  ) +
  labs(
    colour = "Number of Clusters",
    x = "Difference to cluster mean"
  ) +
  theme(
    legend.title.position = "top",
    legend.position = "bottom",
    plot.margin     = margin(2, 2, 2, 2),
    panel.spacing   = unit(2, "pt"),
    strip.text      = element_text(margin = margin(1, 1, 1, 1)),
    legend.box.margin = margin(t = 0, r = 0, b = 0, l = 0),
    legend.margin     = margin(0, 0, 0, 0)
  ) +
  scale_colour_viridis_d(
    option = "plasma", # "turbo",
    # direction = 1,
    name = "Number of Clusters"
  )
p_dens
```
### Figure 2 creation
Creating a plot that shows all 3 maps in one.
```{r}
# ---- Compute common bounding box ----------------------------------------

bboxes <- list(
  st_bbox(df5k_sf),
  st_bbox(df10k_sf),
  st_bbox(df15k_sf)
)

xmin <- min(sapply(bboxes, function(x) x["xmin"]))
xmax <- max(sapply(bboxes, function(x) x["xmax"]))
ymin <- min(sapply(bboxes, function(x) x["ymin"]))
ymax <- max(sapply(bboxes, function(x) x["ymax"]))

coord_common <- coord_sf(
  xlim = c(xmin, xmax),
  ylim = c(ymin, ymax),
  expand = FALSE
)

# ---- Common theme + color scale ----------------------------------------

common_theme <- theme_pubr() +
  theme(
    legend.title.position = "top",
    legend.position = "bottom",
    plot.title = element_text(hjust = 0.5, face = "bold")
  )

color_scale <- scale_color_viridis_c(
  option = "plasma",
  direction = 1,
  name = "Cluster"
)

# ---- Build plots --------------------------------------------------------

df5k_sf$nr_clst <- "5k"
df10k_sf$nr_clst <- "10k"
df15k_sf$nr_clst <- "15k"

# df5_15k <- bind_rows(
#   df5k_sf,
#   df10k_sf,
#   df15k_sf
# ) %>% 
#   mutate(
#     nr_clst = factor(.$nr_clst, levels = c("5k", "10k", "15k"))
#   )

# Choose how many colour groups you can actually see
n_groups <- 12

df5_15k <- bind_rows(
  df5k_sf %>% mutate(nr_clst = "5k"),
  df10k_sf %>% mutate(nr_clst = "10k"),
  df15k_sf %>% mutate(nr_clst = "15k")
) %>%
  mutate(
    nr_clst = factor(nr_clst, levels = c("5k", "10k", "15k")),
    # Stable integer id per original cluster
    cluster_idx = as.integer(factor(cluster)),
    # Modulo grouping (0..n_groups-1 -> then +1 to get 1..n_groups)
    cluster_mod = ((cluster_idx - 1) %% n_groups) + 1,
    cluster_mod = factor(cluster_mod)  # discrete mapping
  )

p_left <- ggplot(df5_15k) +
  facet_wrap(~nr_clst, nrow = 3, ncol = 1) +
  geom_sf(aes(color = cluster_mod), size = 1) + # 0.2) +
  MetBrewer::scale_color_met_d("Redon") + #, guide = "none") +
  coord_common +
  common_theme +
  theme(
    axis.text.x  = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.y  = element_blank(),
    axis.title.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "none", # "right",
    plot.margin = margin(2,2,2,2),
    panel.spacing = unit(2, "pt"),
    strip.text = element_text(margin = margin(1, 1, 1, 1))
  )

# ---- Add anomaly plots -------------------------------------------------
p_dens <- p_dens + common_theme

p_fin <- (p_left | p_dens) +
  plot_annotation(tag_levels = "A") +
  plot_layout(widths = c(1,2)) #& #, guides = "collect") &
  # theme(
  #   legend.position = "bottom",
  #   legend.spacing = unit(1, "pt"),
  #   plot.margin = margin(2,10,2,2)
  # )

# ggsave(
#   filename = file.path(getwd(), "output/tmp_fig2.png"),
#   plot = p_fin,
#   width = 100,
#   height = 200,
#   units = "mm",
#   dpi = 100
# )

p_fin
```

```{r}
ggsave(
  filename = file.path(getwd(), "output/Fig2.tiff"),
  plot = p_fin,
  width = 190,
  height = 280,
  units = "mm",
  dpi = 500
)
```
## Experimenting with map visualization
### Modulo to shift colourspace
```{r}
library(dplyr)
library(ggplot2)
library(MetBrewer)
library(paletteer)
# library(viridisLite)  # just for palette length if you want
# library(viridis)    # if you want scale_* functions

# Choose how many colour groups you can actually see
n_groups <- 12

df5_15k <- bind_rows(
  df5k_sf %>% mutate(nr_clst = "5k"),
  df10k_sf %>% mutate(nr_clst = "10k"),
  df15k_sf %>% mutate(nr_clst = "15k")
) %>%
  mutate(
    nr_clst = factor(nr_clst, levels = c("5k", "10k", "15k")),
    # Stable integer id per original cluster
    cluster_idx = as.integer(factor(cluster)),
    # Modulo grouping (0..n_groups-1 -> then +1 to get 1..n_groups)
    cluster_mod = ((cluster_idx - 1) %% n_groups) + 1,
    cluster_mod = factor(cluster_mod)  # discrete mapping
  )

# # A discrete Viridis scale; use _d for discrete, not _c
# color_scale_discrete <- scale_color_viridis_d(
#   option = "plasma",
#   direction = 1,
#   name = paste0("Cluster (mod ", n_groups, ")")
# )

p_left <- ggplot(df5_15k) +
  facet_wrap(~nr_clst, nrow = 3, ncol = 1) +
  geom_sf(aes(color = cluster_mod), size = 0.2) +
  # color_scale_discrete +
  # scale_color_discrete() +
  # paletteer::scale_color_paletteer_d("dichromat::Categorical_12") +
  # MetBrewer::scale_color_met_d("Signac") +
  MetBrewer::scale_color_met_d("Redon") +
  coord_common +
  common_theme +
  theme(
    axis.text.x  = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.y  = element_blank(),
    axis.title.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "none", # "right",
    plot.margin = margin(2,2,2,2),
    panel.spacing = unit(2, "pt"),
    strip.text = element_text(margin = margin(1, 1, 1, 1))
  )

p_left
```

```{r}
p_fin <- (p_left | p_dens) +
  plot_annotation(tag_levels = "A") +
  plot_layout(widths = c(1,2))

ggsave(
  filename = file.path(getwd(), "output/Fig2.tiff"),
  plot = p_fin,
  width = 190,
  height = 280,
  units = "mm",
  dpi = 500
)
```


## Calculating Metrics - 10k clusters
Metrics to calculate: standard deviations, spatial range, relative variability.
```{r}
# mean an standard deviation
cluster_sums <- df10k_sf %>% 
  st_drop_geometry() %>% 
  group_by(cluster) %>% 
  summarise(
    # central metrics (mean and standard deviation)
    max_temp_mean = mean(max_tmp, na.rm = TRUE),
    max_temp_sd = sd(max_tmp, na.rm = TRUE),
    min_temp_mean = mean(min_tmp, na.rm = TRUE),
    min_temp_sd = sd(min_tmp, na.rm = TRUE),
    prec_mean = mean(prec, na.rm = TRUE),
    prec_sd = sd(prec, na.rm = TRUE),
    rad_mean = mean(rad, na.rm = TRUE),
    rad_sd = sd(rad, na.rm = TRUE),
    vpd_mean = mean(vpd, na.rm = TRUE),
    vpd_sd = sd(vpd, na.rm = TRUE),
    
    # relative variability (coefficient of variation)
    max_temp_cv = sd(max_tmp, na.rm = TRUE) / mean(max_tmp, na.rm = TRUE),
    min_temp_cv = sd(min_tmp, na.rm = TRUE) / mean(min_tmp, na.rm = TRUE),
    prec_cv = sd(prec, na.rm = TRUE) / mean(prec, na.rm = TRUE),
    rad_cv = sd(rad, na.rm = TRUE) / mean(rad, na.rm = TRUE),
    vpd_cv = sd(vpd, na.rm = TRUE) / mean(vpd, na.rm = TRUE)
  ) %>% 
  ungroup()

anomaly_sums <- df10k_sf %>%
  st_drop_geometry() %>% 
  group_by(cluster) %>% 
  summarise(
    # central metrics (mean and standard deviation)
    max_temp_mean = mean(max_temp_anomaly, na.rm = TRUE),
    max_temp_sd = sd(max_temp_anomaly, na.rm = TRUE),
    min_temp_mean = mean(min_temp_anomaly, na.rm = TRUE),
    min_temp_sd = sd(min_temp_anomaly, na.rm = TRUE),
    prec_mean = mean(prec_anomaly, na.rm = TRUE),
    prec_sd = sd(prec_anomaly, na.rm = TRUE),
    rad_mean = mean(rad_anomaly, na.rm = TRUE),
    rad_sd = sd(rad_anomaly, na.rm = TRUE),
    vpd_mean = mean(vpd_anomaly, na.rm = TRUE),
    vpd_sd = sd(vpd_anomaly, na.rm = TRUE),
    
    # relative variability (coefficient of variation)
    max_temp_cv = sd(max_temp_anomaly, na.rm = TRUE) / mean(max_temp_anomaly, na.rm = TRUE),
    min_temp_cv = sd(min_temp_anomaly, na.rm = TRUE) / mean(min_temp_anomaly, na.rm = TRUE),
    prec_cv = sd(prec_anomaly, na.rm = TRUE) / mean(prec_anomaly, na.rm = TRUE),
    rad_cv = sd(rad_anomaly, na.rm = TRUE) / mean(rad_anomaly, na.rm = TRUE),
    vpd_cv = sd(vpd_anomaly, na.rm = TRUE) / mean(vpd_anomaly, na.rm = TRUE)
  ) %>% 
  ungroup()
```
Summary of cluster and anomaly sums.
```{r}
summary(cluster_sums)
summary(anomaly_sums)
```

### Cluster area - Table 3 + Figure 3
Compactness - Isoperimetric Quotient (4piA/P²)
- Range 0 - 1
- 1 = A perfect circle (maximum compactness)
- ~0.785 (pi/4) = a perfect square
- Near 0 = highly elongated, irregular or "spidery" shapes least compact
```{r}
area_stats <- df10k_sf %>% 
  group_by(cluster) %>% 
  summarise(
    # hull = st_convex_hull(st_union(geometry)),
    uni = st_union(geometry),
    area_km2 = round(as.numeric(st_area(uni)) /1e6, 3),
    perimeter_km = st_perimeter(uni) / 1000, # Umfang in [km]
    compactness = round((4 * pi * as.numeric(area_km2)) / (as.numeric(perimeter_km)^2), 3) # Edge to area ratio
  ) %>% 
  ungroup()

summary(area_stats)
# Compactness - Isoperimetric Quotient (4piA/P²)
# Range 0 - 1
# 1 = A perfect circle (maximum compactness)
# ~0.785 (pi/4) = a perfect sqare
# Near 0 = highly elongated, irregular or "spidery" shapes least compact

# Build flextable for area stats
l_area_stats <- area_stats %>%
  select(area_km2, perimeter_km, compactness) %>% 
  st_drop_geometry() %>% 
  mutate(perimeter_km = as.numeric(perimeter_km)) %>% 
  pivot_longer(
    cols = c(area_km2, perimeter_km, compactness),
    names_to   = "parameter",
    values_to  = "value"
  ) %>% 
  group_by(parameter) %>%
  summarise(
    min = suppressWarnings(min(value, na.rm = TRUE)),
    max = suppressWarnings(max(value, na.rm = TRUE)),
    .groups = "drop"
  ) %>% 
  mutate(range = paste0(min, "–", max)) %>% 
  select(parameter, range)

ft_a <- flextable(l_area_stats) %>%
  merge_v(j = "ecosystem") %>%
  theme_booktabs() %>%
  color(i = seq(2, nrow(l_area_stats), 2), color = "black", part = "body") %>%
  bg(i = seq(2, nrow(l_area_stats), 2), bg = "#F9F9F9", part = "body") %>%
  hline(i = c(3)) %>%
  autofit() %>%
  align(align = "center", part = "all") %>%
  bold(part = "header")

ft_a <- fit_to_width(ft_a, max_width = 6.5)  
ft_a

mean(area_stats$area_km2)
sd(area_stats$area_km2)
```

#### Figure 3
Creating figure 3 - calculating some summary statistics and after that plotting.
```{r fig3_summary_stat}
area_stats_summary <- area_stats %>%
  summarise(
    area_med = median(area_km2, na.rm = TRUE),
    area_p05 = quantile(area_km2, 0.05, na.rm = TRUE),
    area_p25 = quantile(area_km2, 0.25, na.rm = TRUE),
    area_p75 = quantile(area_km2, 0.75, na.rm = TRUE),
    area_p95 = quantile(area_km2, 0.95, na.rm = TRUE),

    comp_med = median(compactness, na.rm = TRUE),
    comp_p05 = quantile(compactness, 0.05, na.rm = TRUE),
    comp_p25 = quantile(compactness, 0.25, na.rm = TRUE),
    comp_p75 = quantile(compactness, 0.75, na.rm = TRUE),
    comp_p95 = quantile(compactness, 0.95, na.rm = TRUE),

    per_med = median(as.numeric(perimeter_km), na.rm = TRUE),
    per_p05 = quantile(as.numeric(perimeter_km), 0.05, na.rm = TRUE),
    per_p25 = quantile(as.numeric(perimeter_km), 0.25, na.rm = TRUE),
    per_p75 = quantile(as.numeric(perimeter_km), 0.75, na.rm = TRUE),
    per_p95 = quantile(as.numeric(perimeter_km), 0.95, na.rm = TRUE)
  )
```
Creating the figures and saving them to file.
```{r}
ph_area <- ggplot(data = area_stats, aes(x = area_km2)) +
  ggpubr::theme_pubr() + #theme_minimal() +
  geom_freqpoly(binwidth = 1) +
  # geom_vline(
  #   xintercept = area_stats_summary$area_med,
  #   linetype = "solid", linewidth = 0.7
  # ) +
  labs(x = "area [km²]", title = "Cluster area", y = "count") +
  geom_label(
    data = data.frame(
      x = Inf, y = Inf,
      label = 
        sprintf(
          "Median: %.2f\n05%%: %.2f\n25%%: %.2f\n75%%: %.2f\n95%%: %.2f",
          area_stats_summary$area_med,
          area_stats_summary$area_p05,
          area_stats_summary$area_p25,
          area_stats_summary$area_p75,
          area_stats_summary$area_p95
        )
    ),
    aes(x = x, y = y, label = label),
    inherit.aes = FALSE,
    # Position of box
    hjust = 1.1, vjust = 1.2,
    family = "mono",
    lineheight = 1.1,
    size = 4.5,
    label.size = 0.2
  )


ph_comp <- ggplot(data = area_stats) +
  ggpubr::theme_pubr() + #theme_minimal() +
  geom_freqpoly(
    aes(x = compactness),
    binwidth = 0.01,
  ) +
  # geom_vline(
  #   xintercept = area_stats_summary$comp_med,
  #   linetype = "solid", linewidth = 0.7
  # ) +
  labs(title = "Cluster compactness", y = "count") +
  geom_label(
    data = data.frame(
      x = Inf, y = Inf,
      label = 
        sprintf(
          "Median: %.2f\n05%%: %.2f\n25%%: %.2f\n75%%: %.2f\n95%%: %.2f",
          area_stats_summary$comp_med,
          area_stats_summary$comp_p05,
          area_stats_summary$comp_p25,
          area_stats_summary$comp_p75,
          area_stats_summary$comp_p95
        )
    ),
    aes(x = x, y = y, label = label),
    inherit.aes = FALSE,
    # Position of box
    hjust = 1.1, vjust = 1.2,
    family = "mono",
    lineheight = 1.1,
    size = 4.5,
    label.size = 0.2
  )


ph_per <- ggplot(data = area_stats) +
  ggpubr::theme_pubr() + #theme_minimal() +
  geom_freqpoly(
    aes(x = as.numeric(perimeter_km)),
    binwidth = 10,
  ) +
  # geom_vline(
  #   xintercept = area_stats_summary$per_med,
  #   linetype = "solid", linewidth = 0.7
  # ) +
  labs(x = "perimeter [km]", title = "Cluster perimeter", y = "count") +
  geom_label(
    data = data.frame(
      x = Inf, y = Inf,
      label = 
        sprintf(
          "Median: %.2f\n05%%: %.2f\n25%%: %.2f\n75%%: %.2f\n95%%: %.2f",
          area_stats_summary$per_med,
          area_stats_summary$per_p05,
          area_stats_summary$per_p25,
          area_stats_summary$per_p75,
          area_stats_summary$per_p95
        )
    ),
    aes(x = x, y = y, label = label),
    inherit.aes = FALSE,
    # Position of box
    hjust = 1.1, vjust = 1.2,
    family = "mono",
    lineheight = 1.1,
    size = 4.5,
    label.size = 0.2
  )

ph_area + ph_comp + ph_per + plot_annotation(tag_levels = "A")
# ggsave(
#   filename = file.path(getwd(), "output/fig3_areahisto.png"),
#   scale = 4,
#   width = 50,
#   height = 25,
#   units = "mm"
# )
ggsave(
  filename = file.path(getwd(), "output/Figure3.tiff"),
  scale = 3,
  width = 90,
  height = 45,
  units = "mm",
  dpi = 500
)
```
#### Table 3
combining metrics into a table:
```{r}
# Helper function to format ranges consistently
fmt_num <- function(x, digits = 2) {
  # format numbers with fixed digits, trim unnecessary zeros
  formatC(x, format = "f", digits = digits, drop0trailing = TRUE)
}

# 1) Select only the metric columns (drop cluster id)
metrics_long <- cluster_sums %>%
  select(-cluster) %>%
  # 2) Pivot to long with regex capturing parameter and metric
  #    Column names look like: max_temp_mean, max_temp_sd, max_temp_cv, etc.
  pivot_longer(
    everything(),
    names_to   = c("parameter", "metric"),
    names_pattern = "^(max_temp|min_temp|prec|rad|vpd)_(mean|sd|cv)$",
    values_to  = "value"
  )

# 3) Compute min and max for each parameter × metric across all clusters
minmax_tbl <- metrics_long %>%
  group_by(metric, parameter) %>%
  summarise(
    min = suppressWarnings(min(value, na.rm = TRUE)),
    max = suppressWarnings(max(value, na.rm = TRUE)),
    .groups = "drop"
  ) %>%
  # 4) Build the "min–max" string with nice formatting
  mutate(range = paste0(fmt_num(min), "–", fmt_num(max))) %>%
  select(metric, parameter, range) %>%
  # 5) Wider table: columns are parameters
  pivot_wider(
    names_from  = parameter,
    values_from = range
  ) %>%
  # 6) Nice ordering and labeling for metrics
  mutate(metric = recode(metric,
                         mean = "mean",
                         sd   = "sd",
                         cv   = "relative variability")) %>%
  # Optional: order rows in a conventional order
  mutate(metric = factor(metric, levels = c("mean", "sd", "relative variability"))) %>%
  arrange(metric) %>%
  # Optional: nicer parameter labels (keep if you want short names)
  rename(
    max_temp = max_temp,
    min_temp = min_temp,
    prec     = prec,
    rad      = rad,
    vpd      = vpd
  )

minmax_tbl


mean_range_tbl <- anomaly_sums %>%
  summarise(across(
    .cols = c(max_temp_mean, min_temp_mean, prec_mean, rad_mean, vpd_mean),
    .fns  = list(min = ~min(.x, na.rm = TRUE),
                 max = ~max(.x, na.rm = TRUE))
  )) %>%
  pivot_longer(
    everything(),
    names_to = c("parameter", "metric", ".value"),
    names_pattern = "^(.*)_(.*)_(min|max)$"
  ) %>%
  mutate(
    metric = "mean_range_anomaly",
    range  = paste0(fmt_num(min), "–", fmt_num(max))
  ) %>%
  select(metric, parameter, range) %>%
  pivot_wider(names_from = parameter, values_from = range)

mean_range_tbl

fin_tbl <- bind_rows(minmax_tbl, mean_range_tbl)


ft <- flextable(fin_tbl) %>%
  merge_v(j = "ecosystem") %>%
  theme_booktabs() %>%
  color(i = seq(2, nrow(fin_tbl), 2), color = "black", part = "body") %>%
  bg(i = seq(2, nrow(fin_tbl), 2), bg = "#F9F9F9", part = "body") %>%
  hline(i = c(4)) %>%
  autofit() %>%
  align(align = "center", part = "all") %>%
  bold(part = "header")

ft <- fit_to_width(ft, max_width = 6.5)  
ft
save_as_docx(ft, path  = file.path(getwd(),"output/metrics_tbl3.docx"))
save_as_docx(ft_a, path = file.path(getwd(), "output/area_mtr_tbl.docx"))
```


```{r}
area_stats %>% filter(area_km2 >= 20000) %>% dplyr::select(one_of("cluster")) %>% st_drop_geometry()


ggplot(
  data = df10k_sf %>%
    filter(
      cluster %in% (area_stats %>% filter(area_km2 >= 10000) %>% dplyr::select(one_of("cluster")) %>% st_drop_geometry() %>% dplyr::pull(cluster))
    ) 
) +
  geom_sf(aes(color = cluster), size = 1) +
  color_scale +
  coord_common
  # labs(title = "10k Clusters", x = NULL, y = "Latitude") +
  # common_theme +
  # theme(
  #   axis.text.y  = element_blank(),
  #   axis.title.y = element_blank(),
  #   axis.ticks.y = element_blank()
  # )

ggplot(data = area_stats) +
  geom_sf(aes(fill = cluster))
```

