# Merging current climate data
First libraries:
```{r}
library(tidyverse)
library(terra)
```
Than setting path independent of operating system.
```{r}
if (.Platform$OS.type == "windows") {
  cPath <- "F:/Data/climatedata/germany_1km_res/2_work"
} else {
  cPath <- "/data/public/Data/climatedata/germany_1km_res/2_work"
}
cPath <- normalizePath(cPath, winslash = "/")
```
## 1. Loading Files
Loading two files:
- k1000_centoids.csv (containing multi annual mean clim data for each cluster)
- cluster_coords.csv (latitude and longitute of each cluster centioid)
```{r}
df_clim <- read_csv(
  file = file.path(cPath, "clustered", "k10000_centoids.csv")
)
df_coords <- read_csv(
  file = file.path(cPath, "cluster_coords.csv")
)
df_ger <- read_csv(
  file = file.path(cPath, "clim_dat.csv")
)
df_gerClust <- read_csv(
  file = file.path(cPath, "clustered", "k10000_clim_dat.csv")
)
# df_data <- inner_join(
#   x = df_clim,
#   y = df_coords,
#   by = join_by(cluster_id == cluster)
# ) %>% dplyr::select(-rep_index)

df_data <- left_join(
  x = df_gerClust,
  y = df_clim,
  by = join_by(cluster == cluster_id),
  suffix = c("_point", "_cluster")
)
```
Now the the centoids need to be identified and tagged with a TRUE/FALSE flag.
```{r}
df_coords$centoid <- TRUE
df_data <- left_join(
  x = df_data,
  y = df_coords,
  by = join_by(cluster == cluster, lon, lat)
)
df_data[is.na(df_data$centoid),]$centoid <- FALSE
```
Now that the centoids are marked and all the other data is combined, it can be transfered into a .tiff file and saved.

## 2. Creating a .tiff
```{r}
raster <- terra::rast(
  df_data, # %>% rename(x = lon, y = lat),
  type = "xyz",
  crs = "EPSG:31467"
)
# Plot to check
plot(raster[["centoid"]])
```
## Saving to file
```{r}
terra::writeRaster(
  x = raster,
  filename = file.path("tif","cluster_clim.tif"),
  overwrite = TRUE
)
write_csv(
  x = df_data,
  file = file.path("clustered", "multi_year_mean_climate1981_2010.csv")
)
```
