# Correcting the climate data for location
The current climate data only represents around 2.5k points all over GER, so there needs to be some adjustment so that all of the 10k clusters have their own climate timeline. For this a correction needs to be done. This will be done in roughly these steps:
- Gathering 30 year periods of climate data and calculating a mean (1981-2010). Here the difficulty is that in the modelled climate the historical data ranges from 1980-2005, so the last 5 years need to be taken from each of the 3 climate scenarios (2.6, 4.5, 8.5) and this for all of the 3 climate models.
- Calculating a correction factor for every cluster (10k correction factors)
- Applying the correction factor to the future climate data

*First some initial setup*
```{r}
# Libraies
library(tidyverse)
library(terra)
library(sf)
library(DBI)
library(RSQLite)
library(sparklyr)
library(collapse)
library(progress)
# library(future)
# library(future.apply)
# library(progressr)
# 
# progressr::handlers(global = TRUE)
# progressr::handlers("txtprogressbar") # or "cli" or "progress"

# Paths
if (.Platform$OS.type == "windows") {
  cPath <- "F:/Data/climatedata/germany_1km_res/2_work"
  path <- "F:/Projects/FutureForest/FF_main/04_work/marc/"
  path_dss <- "/data/dss/SVD/data/climate_database/spark_db_v3/"
} else {
  cPath <- "/data/public/Data/climatedata/germany_1km_res/2_work"
  path <- "/data/public/Projects/FutureForest/FF_main/04_work/marc/"
  path_dss <- "/data/dss/SVD/data/climate_database/spark_db_v3/"
}
cPath <- normalizePath(cPath, winslash = "/")
```

# 30 year periods
## Historical DWD data
Importing the climate data from the DWD and filtering them to the centoids of each cluster (so that only the centoids of the cluster are present - 10k rows)
```{r}
dfDWD <- read_csv(
  file = file.path(cPath, "clustered", "multi_year_mean_climate1981_2010.csv")
) %>% filter(centoid == TRUE)
```
Next step is to check which reference point is closest to the centoids, this is due to the resolution of the climate data in Spark is on a 14km² raster, compared to the 1km² from the DWD.
*Creating a function to gather the reference points: Expects a reference grid and table as well as a dataframe that is to be related to these reference data.*
```{r}
g_gaus_krueger <- "EPSG:31467"
g_wgs84 <- "EPSG:4326"
f_getPointID <- function(x, ref_rast) {
  coords <- x %>% dplyr::select(lat, lon)
  n <- nrow(coords)
  
  # Preallocate result vector
  ref_pnt_vec <- rep(NA, n)
  
  pb <- progress::progress_bar$new(
    format = " Processing [:bar] :percent ETA :eta",
    total = n, clear = FALSE, width = 60
  )
  #cat("Start loop.. \n")
  for (i in seq_len(n)) {
    lat <- as.numeric(coords[i, "lat"])
    lon <- as.numeric(coords[i, "lon"])
    #cat("Coord: ", coord)
    # coord <- coords[1,]
    # 1 Transformation
    coord <- st_sfc(st_point(c(lon, lat)), crs = g_gaus_krueger)
    coord <- st_transform(coord, g_wgs84)
    coord <- st_coordinates(coord)
    # coord <- vect(matrix(c(lon, lat), ncol = 2), crs = g_gaus_krueger)
    # coord <- project(coord, g_wgs84)
    # coord <- crds(coord)
    
    # # Transform to raster CRS
    # coord <- vect(matrix(coord, ncol = 2), crs = g_wgs84) %>% 
    #   project(., crs(ref_rast))
    
    # 2 Find closest raster cell
    cell_id <- cellFromXY(ref_rast, coord)
    cell_cnt <- xyFromCell(ref_rast, cell_id)
    ref_id <- terra::extract(ref_rast, cell_cnt)
    
    # 3 if na search nearby
    if (is.na(ref_id[[1]])) {
      #cat("Closest cell is NA, searching nearby...\n"); flush.console()
      
      # Create a buffer around point
      p <- vect(coord, type = "points", crs = g_wgs84)
      p_buff <- buffer(p, width = 28000) # 14 km buffer
      
      # Crop raster to buffer
      cropped_r <- crop(ref_rast, p_buff)
      
      # Get cell IDs, coords, and values
      cell_ids <- which(!is.na(values(cropped_r)))
      if (length(cell_ids) == 0) {
        warning("No valid cells found nearby.")
        ref_id <- NA
      } else {
        ref_coord <- xyFromCell(cropped_r, cell_ids)
        ref_val <- values(cropped_r)[cell_ids]
        
        # Compute distance from point to each valid cell
        ddist <- sqrt((ref_coord[,1] - coord[1,1])^2 + (ref_coord[,2] - coord[1,2])^2)
        closest_idx <- which.min(ddist)
        
        # Update ref_id and center coord
        ref_id <- ref_val[closest_idx]
        cell_cnt <- ref_coord[closest_idx,]
      }
    }
    
    # Compute distance (Eucl in meter)
    ddist <- sqrt(sum((coord - cell_cnt)^2))
    
    # cat("Closest raster cell id: ", cell_id)
    # cat(" with ref_id: ", ref_id[[1]], "\n")
    # cat("Raster center coord: ", cell_cnt, "; Point coord: ", coord, "\n")
    # cat("Distance to center: ", ddist, "m\n")
    # flush.console()
    
    ref_pnt_vec[i] <- ref_id
    pb$tick()
  }
  return(ref_pnt_vec)
}

# f_getPointID <- function(x, rast, ref_tab){
#   # extract the grid ID
#   lon <- x$lon
#   lat <- x$lat
#   
#   coords <- as.data.frame(cbind(lon, lat))
#   point_extr <- terra::extract(rast, coords)[,2]
#   
#   na_indices <- which(is.na(point_extr))
#   if(length(na_indices) > 0) {
#     for (i in na_indices) {
#       lon_i <- lon[i]
#       lat_i <- lat[i]
#       closest_grid <- ref_tab[which.min(abs(lon_i - ref_tab$gk_x) + abs(lat_i - ref_tab$gk_y)),]
#       point_extr[i] <- as.numeric(closest_grid[1])
#     }
#   }
#   
#   return(point_extr)
# }
```
#### Parallel processing function
```{r}
# f_getPointID_parallel <- function(x, ref_rast) {
#   coords <- x[, c("lat", "lon")]
#   
#   # Setup paralell backend
#   plan(multisession, workers = 10)
#   
#   # Function to process one coordinate
#   process_coord <- function(lat, lon) {
#     p <- progressr::progressor()
#     function(lat, lon) {
#       on.exit(p())  # tick after each item
#       
#       # Transform coordinates
#       coord <- vect(matrix(c(lon, lat), ncol = 2), crs = g_gaus_krueger)
#       coord <- project(coord, g_wgs84)
#       coord_mat <- crds(coord)
#       
#       # Find raster cell
#       cell_id <- cellFromXY(ref_rast, coord_mat)
#       cell_cnt <- xyFromCell(ref_rast, cell_id)
#       ref_id <- terra::extract(ref_rast, cell_cnt)[[1]]
#       
#       # Handle missing data
#       if (is.na(ref_id)) {
#         pnt <- vect(coord_mat, type = "points", crs = g_wgs84)
#         p_buff <- buffer(pnt, width = 14000)
#         cropped_r <- crop(ref_rast, p_buff)
#         
#         cell_ids <- which(!is.na(values(cropped_r)))
#         if (length(cell_ids) == 0) return(NA)
#         
#         ref_coord <- xyFromCell(cropped_r, cell_ids)
#         ref_val <- values(cropped_r)[cell_ids]
#         
#         dists <- sqrt((ref_coord[,1] - coord_mat[1,1])^2 + (ref_coord[,2] - coord_mat[1,2])^2)
#         closest_idx <- which.min(dists)
#         
#         return(ref_val[closest_idx])
#       }
#       return(ref_id)
#     }
#   }
#   
#   # Enable progress handling
#   handlers(global = TRUE)
#   p <- progressor(along = 1:nrow(coords))
#   
#   result <- with_progress({
#     future_mapply(
#       FUN = process_coord(coords$lat, coords$lon),
#       lat = coords$lat,
#       lon = coords$lon,
#       future.seed = TRUE
#     )
#   })
#   
#   return(result)
# }
```


*Setting up the reference raster and table:*
```{r}
x_rast <- rast(file.path(path, "clim_data", "reference_grid.tif"))
crs(x_rast) <- "+proj=longlat +datum=WGS84 +no_defs"
x_rast <- project(x_rast, g_wgs84)

# # Project the raster back to Gaus-Krueger
# nx_rast <- terra::project(x_rast, "EPSG:31467")
# 
# ref_tab <- read.csv(file.path(path, "clim_data", "reference_grid_tab.csv"))
# # Transform ref_tab also to Gaus-Krueger
# ref_sf <- st_as_sf(ref_tab, coords = c("wgs_x", "wgs_y"), crs = 4326)
# ref_gk <- st_transform(ref_sf, crs = 31467)
# transformed <- st_coordinates(ref_gk)
# ref_tab$gk_x <- transformed[,1]
# ref_tab$gk_y <- transformed[,2]
# head(ref_tab)
# 
# rm(x_rast, ref_sf, ref_gk, transformed)
```
### Link DWD data to reference raster
Using the function to link the DWD centoids to the next available reference grid.
```{r}
# dfDWD$ref_point <- as.integer(f_getPointID(
#   x = dfDWD,
#   rast = nx_rast,
#   ref_tab = ref_tab
# ))
# head(dfDWD)

dfDWD$ref_point <- as.integer(f_getPointID(
  x = dfDWD,
  ref_rast = x_rast
))
summary(dfDWD)
# dfDWD$ref_point <- f_getPointID_parallel(
#   x = dfDWD,
#   ref_rast = x_rast
# )
```
## Climate model data
Gathering 30 years of data (1981-2010) from Spark.
*First some setup:*
```{r}
scenarios_fut <- c(
  "ICHEC-EC-EARTH_rcp_8_5",
  "ICHEC-EC-EARTH_rcp_4_5",
  "ICHEC-EC-EARTH_rcp_2_6",
  "NCC-NorESM1-M_rcp_8_5",
  "NCC-NorESM1-M_rcp_4_5",
  "NCC-NorESM1-M_rcp_2_6",
  "MPI-M-MPI-ESM-LR_rcp_8_5",
  "MPI-M-MPI-ESM-LR_rcp_4_5",
  "MPI-M-MPI-ESM-LR_rcp_2_6"
)
scenarios_hist <- c(
  "ICHEC-EC-EARTH_historical",
  "NCC-NorESM1-M_historical",
  "MPI-M-MPI-ESM-LR_historical"
)

Sys.setenv("SPARK_HOME" = "/opt/spark")
# Set memory allocation for whole local spark instance
Sys.setenv("SPARK_MEM" = "500G") #13g

# Optionally, ensure JAVA_HOME is also visible to the R session,
# though sparklyr often picks this up from Spark's environment if SPARK_HOME is set.
# Replace with your actual Java home if needed (the one you put in spark-env.sh)
Sys.setenv(JAVA_HOME = "/usr/lib/jvm/java-8-openjdk-amd64") # Or your specific path

# Set driver and executor memory allocation
config <- sparklyr::spark_config()
config$`spark.driver.memory` <- "456G" #"64G" # Driver memory (for R + Spark comms)
config$`sparklyr.shell.driver-memory` <- "456G" #"48G"
config$`spark.executor.memory` <- "456G" #"100G" # Executor (worker) memory
config$`spark.driver.maxResultSize` <- "300G" #"20G" # Max size of collect() results
config$`spark.yarn.executor.memoryOverhead` <- "48G" #"6G" # JVM overhead for executor
config$sparklyr.gateway.port = 8892
config$sparklyr.gateway.start.timeout = 180
config$sparklyr.gateway.connect.timeout = 1
config$`sparklyr.cores.local` <- 50 #4
config$`spark.executor.cores` <- 50 #4
config$`spark.network.timeout` <- "1200s" # 20 min timeout
config$`spark.executor.heartbeatInterval` <- "60s"
config$spark.sql.catalogImplementation <- "in-memory"
config$spark.hadoop.hive.metastore.uris <- ""  # explicitly remove Hive URI

options(java.parameters = "-Xmx456G") # "-Xmx8048m")

f_hive_free_read_paquete <- function(sc, path, tbl_name = NULL) {
  sdf <- spark_session(sc) %>%
    invoke("read") %>%
    invoke("parquet", path)
  
  if (!is.null(tbl_name)) {
    sdf <- sparklyr::sdf_register(sdf, tbl_name)
  }
  
  return(sdf)
}
```
### Data gathering
Functionality of gathering data from spark for each cluster, together with bias correction. First bias factor is calculated and saved into files. Later this bias factor is used to "correct" the data gathered for a point towards the cluster.
```{r}
ref_points <- unique(dfDWD$ref_point)
for (scn in scenarios_hist) {
  years <- c(1981:2005)
  years_fut <- c(2006:2010)
  
  # Connect to spark
  sc <- sparklyr::spark_connect(master = "local", config = config)
  cat("Spark connected..\n"); flush.console()
  
  # Spark handler
  try({
    spark_session(sc) %>%
      invoke("catalog") %>%
      invoke("dropTempView", "climate")
    },
    silent = TRUE
  )
  
  # Read paquetes from spark
  sprk_tbl_handle <- f_hive_free_read_paquete(
    sc = sc,
    path = file.path(path_dss, paste0(scn, "/")),
    tbl_name = "climate"
  )
  cat("Table handler created..\n"); flush.console()
  
  # Gather historic climate data from spark
  clim_tbl <- tryCatch({
    sprk_tbl_handle %>% 
      sparklyr::filter(point_id %in% ref_points) %>% 
      sparklyr::filter(year %in% years) %>% 
      sparklyr::select(point_id, year, month, day, min_temp, max_temp, prec, rad, vpd) %>%
      sparklyr::distinct() %>% 
      sparklyr::collect()
  }, error = function(e) {
    cat("Error during Spark collect: ", e$message, "\n")
    flush.console()
    return(NULL)
  })
  if (is.null(clim_tbl)) {
    next
  }
  clim_fut_batch <- list()
  clim_tbl_out <- clim_tbl %>% 
    mutate(year = as.numeric(year),
           day = as.integer(day))
  cat("Data gathered...\n"); flush.console()
  
  # Gather the few years left from the future scenarios
  for (sp_fut in scenarios_fut) {
    if (strsplit(scn, "_")[[1]][1] != strsplit(sp_fut, "_")[[1]][1]) {
      cat(
        strsplit(scn, "_")[[1]][1],
        " not matching ",
        strsplit(sp_fut, "_")[[1]][1],
        " skipping..\n"
      )
      flush.console()
      next
    }
    # Read parquetes from spark
    sprk_tbl_handle_fut <- f_hive_free_read_paquete(
      sc = sc,
      path = file.path(path_dss, paste0(sp_fut, "/")),
      tbl_name = "climate"
    )
    cat("Future table handler created..\n"); flush.console()
    
    # Gather future climate data from spark
    fut_clim <- tryCatch({
      sprk_tbl_handle_fut %>% 
        sparklyr::filter(point_id %in% ref_points) %>% 
        sparklyr::filter(year %in% years_fut) %>% 
        sparklyr::select(point_id, year, month, day, min_temp, max_temp, prec, rad, vpd) %>% 
        sparklyr::distinct() %>% 
        sparklyr::collect()
    }, error = function(e) {
      cat("Error during Spark collect: ", e$message, "\n")
      flush.console()
      next
    })
    cat("Data gathered...\n"); flush.console()
    clim_fut_batch[[sp_fut]] <- fut_clim
    
    rm(sprk_tbl_handle_fut)
  }
  
  df_clim_fut <- data.table::data.table(do.call(bind_rows, clim_fut_batch))
  fut_clim <- collapse::collap(
    df_clim_fut,
    by = ~ point_id + year + month + day,
    FUN = fmean,
    na.rm = TRUE
  )
  clim_tbl_out <- rbind(clim_tbl_out, fut_clim)
  # Write raw data
  write_csv(
    x = clim_tbl_out,
    file = file.path(cPath, "bias_tbl", paste0("r_",scn,"_fut.csv"))
  )
  # Aggregate to 30 year period
  agg_clim <- clim_tbl_out %>% 
    group_by(point_id, year) %>% 
    summarize(
      min_temp = mean(min_temp),
      max_temp = mean(max_temp),
      precip = sum(prec),
      vpd = mean(vpd),
      rad = sum(rad)
    ) %>% ungroup() %>% group_by(point_id) %>% 
    summarize(
      min_temp = mean(min_temp),
      max_temp = mean(max_temp),
      precip = mean(precip),
      vpd = mean(vpd),
      rad = mean(rad)
    )
  # agg_fut <- agg_clim %>% filter(year > 2005) %>% 
  #   mutate(
  #     precip = precip/3,
  #     rad = rad/3
  #   )
  # agg_clim <- bind_rows(
  #   agg_clim %>% filter(year <= 2005),
  #   agg_fut
  # ) %>%
  #   group_by(point_id) %>% 
  #   summarize(
  #     min_temp = mean(min_temp),
  #     max_temp = mean(max_temp),
  #     precip = mean(precip),
  #     vpd = mean(vpd),
  #     rad = mean(rad)
  #   ) %>% ungroup()
  # Write aggregated data
  write_csv(
    x = agg_clim,
    file = file.path(cPath, "bias_tbl", paste0("agg_", scn, ".csv"))
  )
  
  # Finally
  cat("Disconnecting..\n")
  flush.console()
  
  sparklyr::spark_disconnect(sc)
  sparklyr::spark_disconnect_all()
  sc <- NULL
  sprk_tbl_handle <- NULL
  
  cat("Disconnect successfull.\n")
  flush.console()
}
```

```{r}
# cleanup afte the cathering
rm(agg_clim, clim_fut_batch, clim_tbl, clim_tbl_out, config, df_clim_fut)
rm(fut_clim, nx_rast, ref_tab, ref_points, sc, scenarios_fut)
rm(scenarios_hist, scn, sp_fut, sprk_tbl_handle, years, years_fut, agg_fut)
```

## Calculating bias
Now that all the reference climate values are calculated, the bias factor can be calculated from the difference of the reference climate values to the DWD values. **But** the values of precipitation and radiation are sums over the 30years in the reference tables and mean values in the DWD table, so there needs to be a mean calculation before attempting to create the bias factor.
```{r}
df_agg1 <- read_csv(
  file = file.path(cPath,"bias_tbl","agg_ICHEC-EC-EARTH_historical.csv")
)
df_agg2 <- read_csv(
  file = file.path(cPath,"bias_tbl","agg_MPI-M-MPI-ESM-LR_historical.csv")
)
df_agg3 <- read_csv(
  file = file.path(cPath,"bias_tbl","agg_NCC-NorESM1-M_historical.csv")
)

df_agg1$src <- "ICHEC-EC-EARTH"
df_agg2$src <- "MPI-M-MPI-ESM-LR"
df_agg3$src <- "NCC-NorESM1-M"

dfModels <- bind_rows(df_agg1, df_agg2, df_agg3)
rm(df_agg1, df_agg2, df_agg3)

summary(dfModels)

df_biasii <- tibble(
  lat = numeric(),
  lon = numeric(),
  cluster = numeric(),
  ref_point = numeric(),
  src = character(),
  max_tmp_bias = numeric(),
  min_tmp_bias = numeric(),
  prec_bias = numeric(),
  vpd_bias = numeric(),
  rad_bias = numeric()
)
for (c_src in unique(dfModels$src)) {
  # Calculate bias
  df_bias <- left_join(
    x = dfDWD,
    y = dfModels %>% filter(src == c_src),
    by = join_by(ref_point == point_id),
    suffix = c("dwd","sprk")
  ) %>% mutate(
    max_tmp_bias = max_tmp_point - max_temp,
    min_tmp_bias = min_tmp_point - min_temp,
    prec_bias = prec_point / precip,
    vpd_bias = vpd_point / vpd,
    rad_bias = rad_point / rad
  ) %>% dplyr::select(
    lat, lon, cluster, ref_point, src,
    max_tmp_bias, min_tmp_bias,
    prec_bias, vpd_bias, rad_bias
  )
  df_biasii <- rbind(df_biasii, df_bias)
}
df_bias <- df_biasii
rm(df_biasii)

write_csv(
  x = df_bias %>% filter(src == "ICHEC-EC-EARTH"),
  file = file.path(cPath, "bias_tbl", "ICHEC-EC-EARTH_bias.csv")
)
write_csv(
  x = df_bias %>% filter(src == "MPI-M-MPI-ESM-LR"),
  file = file.path(cPath, "bias_tbl", "MPI-M-MPI-ESM-LR_bias.csv")
)
write_csv(
  x = df_bias %>% filter(src == "NCC-NorESM1-M"),
  file = file.path(cPath, "bias_tbl", "NCC-NorESM1-M_bias.csv")
)
```
--The radiation values seem way to high! DWD has values ranging from 14-48 and aggregated spark values range from 3060-6028. So there has to be something missing.--
**Issue fixed**, Caused by wrong DWD rasters for radiation. This was fixed by loading the radiation for every month , creating a a sum and fixing the difference in SI Unit (kWh/m² vs. MJ/m²/day).